---
title: "codebook.md"
author: "Dinesh Choudhari"
date: "February 28, 2016"
output: html_document
---
Information and Description of HCI data set.
==
Human Activity Recognition Using Smartphones Dataset Version 1.0


The dataset includes the following files:

- 'README.txt'
- 'features_info.txt': Shows information about the variables used on the feature vector.
- 'features.txt': List of all features.
- 'activity_labels.txt': Links the class labels with their activity name.
- 'train/X_train.txt': Training set.
- 'train/y_train.txt': Training labels.
- 'test/X_test.txt': Test set.
- 'test/y_test.txt': Test labels.

The following files are available for the train and test data. Their descriptions are equivalent. 

- 'train/subject_train.txt': Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 


PROJECT IMPLEMENTATION:
==

##Step 1: Getting and understanding the source data.

1. Create a new directory HARdata (HumanActivityRecognitionData) if not exists to store data files. Download the file directly from given URL

```{r}
   setwd("~/Desktop/Coursera")
```

```{r}  
   if(!file.exists("./HARdata")){dir.create("./HARdata")}
   fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
   download.file(fileUrl,destfile="./HARdata/Dataset.zip",method="curl")
```


2. Unzip the file and extract files from UCI HAR dataset folder

```{r}
   unzip(zipfile="./HARdata/Dataset.zip",exdir="./HARdata")
   path_rf <- file.path("./HARdata" , "UCI HAR Dataset")
   files<-list.files(path_rf, recursive=TRUE)
```

3. The files that will be used to load data are listed as follows:
 
   test/subject_test.txt 
   test/X_test.txt 
   test/y_test.txt 
   train/subject_train.txt
   train/X_train.txt 
   train/y_train.txt

4. Possible variables based on the file names is
 
  a. Acitvity --> from Y_train.txt and Y_test.txt 
  b. Subject  --> from subject_train.txt and subject_test.txt
  c. Features --> from X_train.txt and X_test.txt
  d. Features --> from features.txt
  e. Acitvity --> from activity_lables.txt
  
  Descriptive variables ( data frame )
  a. Activity
  b. Subject
  c. Features
  
##Step 2: Read and store the files into above descriptive variables (Note the variables will be of structure data.frame)

1. Test and Train Activity files

```{r}
   dataActivityTest  <- read.table(file.path(path_rf, "test" , "Y_test.txt" ),header = FALSE)
   dataActivityTrain <- read.table(file.path(path_rf, "train", "Y_train.txt"),header = FALSE)
```

2. Test and Train subject files.

```{r}
   dataSubjectTrain <- read.table(file.path(path_rf, "train", "subject_train.txt"),header = FALSE)
   dataSubjectTest  <- read.table(file.path(path_rf, "test" , "subject_test.txt"),header = FALSE)
```

3. Test and Train Feature files.

```{r}
   dataFeaturesTest  <- read.table(file.path(path_rf, "test" , "X_test.txt" ),header = FALSE)
   dataFeaturesTrain <- read.table(file.path(path_rf, "train", "X_train.txt"),header = FALSE)
```


##Step 3: R Script run_analysis.R 

1. Merges the training and the test sets to create one data set.
  
 a. R bind data tables

```{r} 
    dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
    dataActivity<- rbind(dataActivityTrain, dataActivityTest)
    dataFeatures<- rbind(dataFeaturesTrain, dataFeaturesTest)
```

 b. Variables names.

```{r} 
    names(dataSubject)<-c("subject")
    names(dataActivity)<- c("activity")
    dataFeaturesNames <- read.table(file.path(path_rf, "features.txt"),head=FALSE)
    names(dataFeatures)<- dataFeaturesNames$V2
```

 c. Column bind to merge for data frame.

```{r} 
    dataCombine <- cbind(dataSubject, dataActivity)
    Data <- cbind(dataFeatures, dataCombine)
```
 

2. Extracts only the measurements on the mean and standard deviation for each measurement.

 a. Take Names of Features with mean() and std()

```{r}    
    subdataFeaturesNames<-dataFeaturesNames$V2[grep("mean\\(\\)|std\\(\\)", dataFeaturesNames$V2)]
```

 b. Data frame data by selected names fo Feature from above a.

```{r}    
    selectedNames<-c(as.character(subdataFeaturesNames), "subject", "activity" )
    Data<-subset(Data,select=selectedNames)
```

    
3. Uses descriptive activity names to name the activities in the data set

 a. Descriptive activity names from “activity_labels.txt”

```{r} 
    activityLabels <- read.table(file.path(path_rf, "activity_labels.txt"),header = FALSE)
```


4. Appropriately labels the data set with descriptive variable names.

   Replace,
-   t    --> time
-   f    --> frequency
-   Gyro --> Gyroscope
-   Mag  --> Magnitude
-   Body --> Body

```{r}   
  names(Data)<-gsub("^t", "time", names(Data))
  names(Data)<-gsub("^f", "frequency", names(Data))
  names(Data)<-gsub("Acc", "Accelerometer", names(Data))
  names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
  names(Data)<-gsub("Mag", "Magnitude", names(Data))
  names(Data)<-gsub("BodyBody", "Body", names(Data))
```

   
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

  a. Tidy data

```{r}    
    library(plyr);
    Data2<-aggregate(. ~subject + activity, Data, mean) 
    Data2<-Data2[order(Data2$subject,Data2$activity),] 
```

   b. Write second tidy data set to file tidydata.txt

```{r}   
      write.table(Data2, file = "tidydata.txt",row.name=FALSE)
```
